Report on Late Winter2013 Production: Image Differencing\vspace{-2ex}

### 2013-04-01 W13report.tex
Working on discussion in 5.1

Detection on the difference image occurs through correlation of
$D(x,y)$ with the science image's PSF, yielding optimally filtered
detection image $D'(x,y) = D(x,y) \circ PSF_S(u,v;x,y)$ where $\circ$
denotes correlation (currently the DM stack uses convolution instead of correlation,
reported as ticket #2552).  The values of the pixels in $D'(x,y)$ provide a
maximum likelihood estimate of there being a point source at that
position.  Detection occurs by simply finding pixels that are more
than N sigma above the square root of the per--pixel variance.

@@

We can model source free image (i.e.\ with a background described by a
Gaussian process) using the statistics of Gaussian random fields
(Kaiser 2004). For an image with Gaussian noise convolved with a
Gaussian PSF with width $\sigma_g$ the density of peaks above a given
detection threshold, $\nu$ is given by 
\begin{equation}
n(>\nu) = \frac{1}{2^{5/2}\pi^{3/2}\sigma_g^{2}} \nu e^{-\nu^2 /2} 
\label{eq-theory}
\end{equation}
 To estimate the number of random
positive detections for a given image we must normalize this density
by the total number of realizations of the PSF within the image, i.e.,
\begin{equation}
N_{total}(>\nu) = n(\nu)*nrow*ncol/PSF_{FWHM}^2
\end{equation}
where nrow and ncol are the number of rows and columns in an images
and $PSF_{FWHM}$ is the FWHM of the PSF.

Figure~\ref{fig:peaks} shows good agreement in a comparison of this
model to a Monte-Carlo simulation for thresholds $4<\nu<6$ for a
4000x4072 pixel image with 0.88 arcsec seeing. The variance on these
measures is shown in Figure XXX {\bf Simon: I can make this figure on Monday morn, if needed.}
\begin{figure}
\includegraphics[width=0.5\textwidth,
height=0.25\textwidth]{figures/peaks.eps} 
\caption{A comparison of the predicted number of random peaks (solid
  line) detected in an image, as a function of significance, with
  those derived from a Monte-Carlo realization of a Gaussian noise
  image (solid points).}
\label{fig:peaks}
\end{figure}

@@

\begin{figure}%
\captionsetup[subfloat]{labelformat=empty}
\centering
\subfloat[]{\includegraphics[width=0.3\textwidth]{figures/fp_v_thresh_01_nocorr.eps}}
\subfloat[]{\includegraphics[width=0.3\textwidth]{figures/fp_v_thresh_02_nocorr.eps}}
\subfloat[]{\includegraphics[width=0.3\textwidth]{figures/fp_v_thresh_03_nocorr.eps}}\\
\subfloat[]{\includegraphics[width=0.3\textwidth]{figures/fp_v_thresh_01_corr.eps}}%
\subfloat[]{\includegraphics[width=0.3\textwidth]{figures/fp_v_thresh_02_corr.eps}}%
\subfloat[]{\includegraphics[width=0.3\textwidth]{figures/fp_v_thresh_03_corr.eps}}%
\caption{Here we show the comparison of detected false positive count to the analytic prediction.
In all cases the blue circles are the prefilter case and the red squares are the postfilter case.  
The top row has had the counts corrected for the underestimate of the noise by the variance plane (Table~\ref{tab-variance1}).  
In the bottom row we undertaken a 1--parameter correction to align the predicted and measured values.  See Section~\ref{sec-thresh}
for how this correction is applied.  The 2.5\%, 1.5\% and 1\% scaling in $k$ correspond to a 5\%, 3\%, 2\% over-estimation of the image variance, respectively
from Left to Right in the bottom pane.  
The shaded area is the 1-$\sigma$ confidence envelope
determined from Monte Carlo simulations assuming the analytic form of the false detection rate and a Gaussian distribution.}
\label{fig-fp_v_thresh}
\end{figure}

>>>>

Detection on the difference image occurs through correlation of
$D(x,y)$ with the science image's PSF, yielding optimally filtered
detection image $D'(x,y) = D(x,y) \circ PSF_S(u,v;x,y)$ where $\circ$
denotes correlation (currently the DM stack uses convolution instead of correlation,
reported as ticket \#2552).  The values of the pixels in $D'(x,y)$ provide a
maximum likelihood estimate of there being a point source at that
position.  Detection occurs by simply finding pixels that are more
than N sigma above the square root of the per--pixel variance.

@@

We model a source--free image (i.e.\ with a background described by a
Gaussian process) using the statistics of Gaussian random fields \citep{Kaiser-PointSources}.  
For an image with Gaussian noise convolved with a
Gaussian PSF with width $\sigma_g$ the density of peaks above a given
detection threshold, $\nu$ is given by 
\begin{equation}
n(>\nu) = \frac{1}{2^{5/2}\pi^{3/2}\sigma_g^{2}} \nu e^{-\nu^2 /2} 
\label{eq-theory}
\end{equation}
 To estimate the number of random
positive detections for a given image we must normalize this density
by the total number of realizations of the PSF within the image, i.e.,
\begin{equation}
N_{total}(>\nu) = n(\nu)*nrow*ncol/PSF_{FWHM}^2
\end{equation}
where nrow and ncol are the number of rows and columns in an images
and $PSF_{FWHM}$ is the FWHM of the PSF.  In difference image analysis, we expect
twice this number, since we are sensitive to both positive and negative--going
fluctuations.

Figure~\ref{fig:peaks} shows good agreement in a comparison of this
model to a Monte-Carlo simulation for thresholds $4<\nu<6$ for a
4000x4072 pixel image with 0.88 arcsec seeing. The variance on these
measures is calculated using 1e6 random realizations.
\begin{figure}
\centering{\includegraphics[width=0.5\textwidth]{figures/peaks.eps}}
\caption{A comparison of the predicted number of random peaks (solid
  line) detected in an image, as a function of significance, with
  those derived from a Monte-Carlo realization of a Gaussian noise
  image (solid points).}
\label{fig:peaks}
\end{figure}

@@

\begin{figure}%
\captionsetup[subfloat]{labelformat=empty}
\centering
\subfloat[]{\includegraphics[width=0.3\textwidth]{figures/fp_v_thresh_01_nocorr.eps}}
\subfloat[]{\includegraphics[width=0.3\textwidth]{figures/fp_v_thresh_02_nocorr.eps}}
\subfloat[]{\includegraphics[width=0.3\textwidth]{figures/fp_v_thresh_03_nocorr.eps}}\\
\subfloat[]{\includegraphics[width=0.3\textwidth]{figures/fp_v_thresh_01_corr.eps}}%
\subfloat[]{\includegraphics[width=0.3\textwidth]{figures/fp_v_thresh_02_corr.eps}}%
\subfloat[]{\includegraphics[width=0.3\textwidth]{figures/fp_v_thresh_03_corr.eps}}%
\caption{Here we show the comparison of detected false positive count to the analytic prediction.
In all cases the blue circles are the prefilter case and the red squares are the postfilter case.  
The top row has had the counts corrected for the underestimate of the noise by the variance plane (Table~\ref{tab-variance1}).  
In the bottom row we undertaken a 1--parameter correction to align the predicted and measured values.  See Section~\ref{sec-thresh}
for how this correction is applied.  The 2.5\%, 1.5\% and 1\% scaling in $k$ correspond to a 5\%, 3\%, 2\% over-estimation of the image variance, respectively
from Left to Right in the bottom pane.  
The shaded area is the 1-$\sigma$ confidence envelope
determined from Monte Carlo simulations (Section~\ref{sec-analyticfp}).}
\label{fig-fp_v_thresh}
\end{figure}
